---
title: "Statistical Analysis of Lana's dataset"
subtitle: "v7 - 2020-09-11"
output:
  html_document:
    df_print: paged
    toc: true
  pdf_document: default
  word_document: default

---

# Install

To download all needed packages, execute renv::restore() the first time:

```{r include=FALSE}
# Packages to install.
#install.packages(c('faraway', 'GGally', 'readxl', 'sjlabelled', 'sjmisc', 'sjPlot', 'tidyverse'))

library(tidyverse)
library(readxl)
library(sjPlot)
library(sjmisc)
library(sjlabelled)

```



# General notes.

- I call all the IV predictors here. They are in %. i.e. 0.1 = 10%

- R2 describes the goodness of fit and how much of the data can be fitted by the regression hyperplane. To compensate for different number of predictors, we only use R_adj to compare different models (of different size). Attention: R2 does not say if any predictors explain the data in the model. We can have quite a good R2 but no significant predictors, or predictors with small effects. In the non-significant cases, we can probably still draw a conclusion for the current sample though (but not generalise)

- For a good model, there must no be any large collinearities between variables. We test collinearities with VIF. values above 5 are critical and above 10 dangerous. We need to test for every model.

- The selection of variables is essential. The preferred way is to have some theorethical knowledge from other studies. However, because we do exploratory research,we apply a stepwise backward selectin of variables. WARGNING: According to my stats script, "There is a general rule of the thumb, saying that there should be at least 5 times as many observations as the number of coefficients that are estimated.". I.e. based on the 8 IV, there should be 40 observations, however, we only have 18 observations.

## Assumptions
We produce the residual plots. From top left to bottom right:

- Plot 1: The average error shoud be 0 (within random resampled lines). Then there is no systematic error.
- Plot 2: Redsiduals must be normally distributed, i.e. the points should lie on the diagonal line
- Plot 3: There should be homoscedasticity, i.e. the line should be straight.
- Plot 4: There should not be influenctial cases. outside the red dotted lines.

If all statistical assumptions are fulfilled, the generalisation is possible and estimation with MLR are robust. If not, google solutions.


```{r load data (from env), echo = FALSE}

load(file = "environment_20200911.RData")
rm(list=setdiff(ls(), "combined_analysis"))

gloabl_show_analytics <- FALSE

```

```{r load functions, include=FALSE}
source("functions.R")

full_models <- list()
full_models[["RE"]] <- list()
full_models[["EE"]] <- list()
full_models[["RE_perf_abs"]] <- list()
full_models[["EE_perf_abs"]] <- list()
full_models[["RE_amb"]] <- list()
full_models[["EE_amb"]] <- list()

formulas <- list()

```

# Data

## Data view

Let's have a look at the original data. We remove several variables that, from a theortical consideration, are not relevant (and would sometimes cause collinearity). We also transform some values to a standard % format. 

```{r dataview}
#summary(combined_analysis)

combined_analysis_transformed <- combined_analysis %>% 
  mutate(
    pre_potential_EI = pre_potential_EI / 100)

summary(combined_analysis_transformed)

```


```{r translation tables, include=FALSE}
# Here we change the column names and prepare the dataframes. Assignements can now be adjusted in the Excel file.

all_variables_translation_table <- read_excel(path = "MLR_variables.xlsx")

# Select all variables in col_names_for_RE	col_names_for_EE, and prepare for renaming

all_variables <- list()
traslation_tables <- list()

# Select RE
all_variables[["RE"]] <- all_variables_translation_table %>% 
  filter(!is.na(col_names_for_RE)) %>% 
  select(col_name = col_names_for_RE, symbol)

traslation_tables[["RE"]] <- all_variables[["RE"]]$col_name
names(traslation_tables[["RE"]]) <- all_variables[["RE"]]$symbol

# Select EE
all_variables[["EE"]] <- all_variables_translation_table %>% 
  filter(!is.na(col_names_for_EE)) %>% 
  select(col_name = col_names_for_EE, symbol)

traslation_tables[["EE"]] <- all_variables[["EE"]]$col_name
names(traslation_tables[["EE"]]) <- all_variables[["EE"]]$symbol


## Gather and rename
                                           
combined_analysis_gathered <- combined_analysis_transformed %>% 
  select(-prioritization_scale, -Prio_EE_REN) %>% 
  gather(key = "variable", value = "value", one_of(all_variables$RE$col_name, all_variables$EE$col_name)) 

Ren_both_effects <- list()
Ren_both_effects[["long"]] <- combined_analysis_gathered %>% 
  filter(variable %in% c(all_variables$RE$col_name)) %>% #remove variables that are not in MLR_variables.xls
  mutate(variable = fct_recode(variable, !!!traslation_tables$RE))

Ren_both_effects[["wide"]] <- Ren_both_effects[["long"]]  %>% 
  select(country_code, variable, value) %>% 
  spread(key = "variable", value = "value")

EE_both_effects <- list()
EE_both_effects[["long"]]  <- combined_analysis_gathered %>% 
  filter(variable %in% c(all_variables$EE$col_name)) %>% #remove variables that are not in MLR_variables.xls 
  mutate(variable = fct_recode(variable, !!!traslation_tables$EE))

EE_both_effects[["wide"]] <- EE_both_effects[["long"]]  %>% 
  select(country_code, variable, value) %>% 
  spread(key = "variable", value = "value")


all_variables_translation_table %>% 
  select(symbol, variable)

variable_names <- all_variables_translation_table$variable
names(variable_names) <- all_variables_translation_table$symbol


```

## Remove missing cases

First we need to remove missing cases.

```{r echo=FALSE}

# View(combined_analysis)

# From here on use wide format due to R methods.
 
EE_both_effects[["wide_no_na"]] <- na.omit(EE_both_effects[["wide"]])
paste("For EE data: Of", nrow(EE_both_effects$wide),"cases", nrow(EE_both_effects$wide_no_na),"are left due to missing data,")
# let's take this sample

Ren_both_effects[["wide_no_na"]] <- na.omit(Ren_both_effects[["wide"]]) %>% 
  filter(country_code %in% EE_both_effects[["wide_no_na"]]$country_code)

paste("For REN data: Of", nrow(Ren_both_effects$wide),"cases", nrow(Ren_both_effects$wide_no_na),"are left due to missing data,")


tibble(
  countries = EE_both_effects[["wide"]]$country_code,
  in_sample = EE_both_effects[["wide"]]$country_code %in% EE_both_effects[["wide_no_na"]]$country_code
) %>% 
  mutate(
    in_sample = if_else(in_sample, "In the finale sample", "Missing")
  ) %>% 
  tab_df()


# Subsample for checking the robustness of the model

ggplot(data = EE_both_effects[["wide_no_na"]], mapping = aes(x = S_size_services, y = EE_perf)) +
  geom_point() + 
  geom_smooth(method = "lm")
```


Some data is missing that we can't use in MLR, but most is there.

## Marginal plots

Let's have a look at the data using marginal plots. This helps to see if any (potential) predictores are skewed. If this is the case, log() transformation will be applied in the regression model, which is a standard procedure to improve the model.

```{r marginal plot, echo=FALSE, message=FALSE, warning=FALSE}

Ren_both_effects %>%
  pluck("long") %>% 
  ggplot(mapping = aes(x = value)) +
    geom_histogram()+
    facet_wrap("variable", scales = "free") +
    labs(title = "Marginal plots for REN data")


EE_both_effects %>%
  pluck("long") %>% 
  ggplot(mapping = aes(x = value)) +
    geom_histogram()+
    facet_wrap("variable", scales = "free") +
    labs(title = "Marginal plots for EE data")

```

```{r variable transformation}

Ren_both_effects[["wide_no_na"]] <- Ren_both_effects[["wide_no_na"]]
EE_both_effects[["wide_no_na"]] <- EE_both_effects[["wide_no_na"]]

```

## Correlations

```{r, warning=FALSE, echo=FALSE}
GGally::ggcorr(Ren_both_effects$wide_no_na, nbreaks = 5, label = TRUE, size = 2, label_size = 3, legend.position = "bottom", legend.size = 10)
```

```{r, warning=FALSE}
GGally::ggcorr(EE_both_effects$wide_no_na %>% select(-EE_perf), nbreaks = 5, label = TRUE, size = 2,label_size = 3, legend.position = "bottom", legend.size = 15)  +
  ggtitle ("Policy & socio-economic effects on EE progress")

```
# Results

## 1) REN Performance (relative)

### Full model

```{r}

full_models[["RE"]][["01_untransformed"]] <- lm(
  RE_perf ~  
    P_density + P_ambition + P_REN +
    S_GAV + S_potential + S_effort + S_PPS + S_elec  + S_centr + S_GDP
  , data = Ren_both_effects[["wide_no_na"]])

summary(full_models$RE$`01_untransformed`)
faraway::vif(full_models$RE$`01_untransformed`)

```

### Stepwise selection

```{r}
full_models[["RE"]][["03_selected"]] <- select_model_and_print_analytics(
  full_models[["RE"]][["01_untransformed"]])

```

Assumptions look ok

### Custom model

```{r}

# Copy formula from full or stepwise model and customise 
full_models[["RE"]][["04_custom"]] <- lm(
  RE_perf ~  
    P_density + P_ambition + P_REN +
    S_GAV + S_potential + S_effort + S_PPS + S_elec  + S_centr + S_GDP,
  Ren_both_effects[["wide_no_na"]]
  )

custom_model_analytics(full_models[["RE"]][["04_custom"]])

```




## 2) REN Performance (absolute)

### Full model
```{r}
full_models[["RE_perf_abs"]][["01_untransformed"]] <- lm(
  
  RE_perf_abs ~ 
    P_density + P_ambition + P_REN +
    S_GAV + S_potential + S_effort + S_PPS + S_elec  + S_centr + S_GDP, 
  
    data = Ren_both_effects[["wide_no_na"]])

summary(full_models$RE$`01_untransformed`)
#faraway::vif(full_models$RE$`01_untransformed`)


```

### Stepwise
```{r}
full_models[["RE_perf_abs"]][["03_selected"]] <- select_model_and_print_analytics(
  full_models[["RE_perf_abs"]][["01_untransformed"]])
```

Looks ok.

### Custom

```{r}
# Copy formula from full or stepwise model and customise 
full_models[["RE_perf_abs"]][["04_custom"]] <- lm(
  RE_perf_abs ~  
    P_density + P_ambition + P_REN +
    S_GAV + S_potential + S_effort + S_PPS + S_elec  + S_centr + S_GDP,
  Ren_both_effects[["wide_no_na"]]
  )

custom_model_analytics(full_models[["RE_perf_abs"]][["04_custom"]])
```


## 3) REN Ambition

```{r}

full_models[["RE_amb"]][["01_untransformed"]] <- lm(
  P_ambition ~ 
    P_density + P_REN + 
    S_GAV + S_potential + S_effort + S_PPS + S_elec + S_centr  + S_GDP, 
  data = Ren_both_effects[["wide_no_na"]])

#filter(country_code != "HR")
#summary(full_models$RE_amb$`01_untransformed`)
#faraway::vif(full_models$RE_amb$`01_untransformed`)

```

```{r}
full_models[["RE_amb"]][["03_selected"]] <- select_model_and_print_analytics(full_models[["RE_amb"]][["01_untransformed"]])
```

Assumptions of homoscedasticity are violated, and there is an influential outlier

```{r}
warning("Check assumptions")
check_outlier(11, data = "RE")
```

### Custom

```{r}
# Copy formula from full or stepwise model and customise 
full_models[["RE_amb"]][["04_custom"]] <- lm(
P_ambition ~ 
    P_density + P_REN + 
    S_GAV + S_potential + S_effort + S_PPS + S_elec + S_centr  + S_GDP,
  Ren_both_effects[["wide_no_na"]]
  )

custom_model_analytics(full_models[["RE_amb"]][["04_custom"]])

```

## 4) EE Performance (relative)

```{r}

full_models[["EE"]][["01_untransformed"]] <- lm(
  EE_perf ~  
    P_density + P_ambition + P_REN +
    S_GAV + S_potential + S_effort + S_PPS + S_elec + S_centr  + S_GDP
  , data = EE_both_effects[["wide_no_na"]])

#filter(!(country_code %in% c("RO","EE"))
#summary(full_models$EE$`01_untransformed`)
#faraway::vif(full_models$EE$`01_untransformed`)


```



```{r}
full_models[["EE"]][["03_selected"]] <- select_model_and_print_analytics(full_models[["EE"]][["01_untransformed"]])
```
Assumptions critical, but iinside resamples area, so acceptable.


```{r}
check_outlier(c(16), "EE")
```


### Custom
(same results as in v4)

```{r}

# Copy formula from full or stepwise model and customise 
full_models[["EE"]][["04_custom"]] <- lm(
  EE_perf ~ P_ambition + S_GAV + S_potential + S_effort + S_PPS,
  EE_both_effects[["wide_no_na"]]
  )

custom_model_analytics(full_models[["EE"]][["04_custom"]])

```

## 5) EE Performance absolute

```{r}
full_models[["EE_perf_abs"]][["01_untransformed"]] <- lm(
  
  EE_perf_abs ~ 
    P_density + P_ambition + P_REN +
    S_GAV + S_potential + S_effort + S_PPS + S_elec  + S_centr + S_GDP, 
  
    data = EE_both_effects[["wide_no_na"]])

#summary(full_models$EE_perf_abs$`01_untransformed`)
##faraway::vif(full_models$EE_perf_abs$`01_untransformed`)


```


```{r}

full_models[["EE_perf_abs"]][["03_selected"]] <- select_model_and_print_analytics(
  full_models[["EE_perf_abs"]][["01_untransformed"]])

```

Assumptions may be slightly violated.

### Custom

```{r}

# Copy formula from full or stepwise model and customise 
full_models[["EE_perf_abs"]][["04_custom"]] <- lm(
  EE_perf_abs ~ 
    P_density + P_ambition + P_REN +
    S_GAV + S_potential + S_effort + S_PPS + S_elec  + S_centr + S_GDP,
  EE_both_effects[["wide_no_na"]]
  )

custom_model_analytics(full_models[["EE_perf_abs"]][["04_custom"]])

```

## 6) EE Ambition


```{r}

full_models[["EE_amb"]][["01_untransformed"]] <- lm(
  P_ambition ~ 
    P_density + P_REN + 
    S_GAV + S_potential + S_effort + S_PPS + S_elec + S_centr  + S_GDP, 
  data = EE_both_effects[["wide_no_na"]])

#summary(full_models$EE_amb$`01_untransformed`)
```


```{r}
full_models[["EE_amb"]][["03_selected"]] <- select_model_and_print_analytics(full_models[["EE_amb"]][["01_untransformed"]])
```

Assumptions slightly skewed, particularly 0-error assumptions, however withing resampled area. So acceptable.

### Custom

```{r}

# Copy formula from full or stepwise model and customise 
full_models[["EE_amb"]][["04_custom"]] <- lm(
  P_ambition ~ 
    P_density + P_REN + 
    S_GAV + S_potential + S_effort + S_PPS + S_elec + S_centr  + S_GDP,
  EE_both_effects[["wide_no_na"]]
  )

custom_model_analytics(full_models[["EE_amb"]][["04_custom"]])

```



# Result tables

```{r}
print_table_of_models <- function(variable, title = ""){
  
  tab_model(
    full_models[[variable]][["01_untransformed"]], 
    full_models[[variable]][["03_selected"]],
    full_models[[variable]][["04_custom"]],

    dv.labels = c("Full model", "Selected model", "Custom"),
    pred.labels = variable_names,
    title = title,
    p.style = "stars",
    show.std = TRUE,
    p.threshold = c(0.1, 0.05, 0.01),
    digits = 3)
  
}

```

```{r}

print_table_of_models("RE", "Explaining REN Performance (relative)")
print_table_of_models("RE_perf_abs", "Explaining REN Performance (absolute)")
print_table_of_models("RE_amb", "Explaining REN Ambition")

print_table_of_models("EE", "Explaining EE Performance (relative)")
print_table_of_models("EE_perf_abs", "Explaining EE Performance (absolute)")
print_table_of_models("EE_amb", "Explaining EE Ambition")

```