---
title: "Statistical Analysis of Lana's dataset"
subtitle: "v2"
output:
  html_document:
    df_print: paged
    toc: true
  pdf_document: default
  word_document: default
---

Second version with additional variables

# General notes.

- I call all the IV predictors here. They are in %. i.e. 0.1 = 10%
- R2 describes the goodness of fit and how much of the data can be fitted by the regression hyperplane. To compensate for different number of predictors, we only use R_adj to compare different models (of different size). Attention: R2 does not say if any predictors explain the data in the model. We can have quite a good R2 but no significant predictors, or predictors with small effects. In the non-significant cases, we can probably still draw a conclusion for the current sample though (but not generalise)
- The correlations do not say anything about the goodnes of fit of the model in MLR.


```{r load, include=FALSE}

#https://briatte.github.io/ggcorr/
#a measure of goodness of fit in multiple regression. -> WRONG (unless I'm wrong :-P9. Where do you have that infofrom?


library(tidyverse)
library(ggrepel)
library(rlang)
library(countrycode) #check
library(rrr)
library(readxl)
library(ggfortify)
library(faraway)
library(broom)
library(sjPlot)
library(sjmisc)
library(sjlabelled)


```


```{r load data, eval=FALSE, include=FALSE, echo = FALSE}
# saved in env
Progress_REN_EE <- read_csv("data_EE_REN.csv")

pre_potential <- read_csv("EE_REN_preexistingpotential.csv")%>%
  rename (pre_potential_EI = "EI_0510")%>%
  rename (pre_potential_REN = "REN_2005_percent")

View(pre_potential)

wealth <- read_csv("EE_REN_wealth.csv")

density <- read_csv("EE_REN_density.csv") %>%
  rename (Policy_density_EE = "percent_EE")%>%
  rename (Policy_density_REN = "percent_REN")

subsidies <- read_csv("EE_REN_subsidies.csv")%>%
  rename (subsidies_REN = "percent_REN")%>%
  rename (subsidies_savings = "percent_savings")

jobs <- read_csv("EE_REN_jobs.csv") %>%
  rename (Jobs_EE = "mean_percent_EE")%>%
  rename (Jobs_REN = "mean_percent_REN")

Tech_potential <- read_csv("EE_REN_potential.csv")%>%
  rename (Jobs_REN = "mean_percent_REN")
View(Tech_potential)

Policy_type <- read_csv("EE_REN_policytype.csv")

Policy_targets <- read_csv("targets_EE_REN.csv")%>%
  select (Country, NREAP_target, target_FEC_renewed)

#combining for analysis


```

```{r load data (from env), echo = FALSE}
#load(file = "Regression_Environment_II.RData")
#load(file = "Data_Paper II.RData")

#load("Environment_20200714.RData")
load(file = "Environment_20200730.RData")


# remmove all but the combined_analysis needed here
rm(list=setdiff(ls(), "combined_analysis"))

```

```{r load functions, include=FALSE}
source("functions.R")

partial_models <- list()
partial_models[["RE"]] <- list()
partial_models[["EE"]] <- list()
partial_models[["RE_amb"]] <- list()
partial_models[["EE_amb"]] <- list()

full_models <- list()
full_models[["RE"]] <- list()
full_models[["EE"]] <- list()
full_models[["RE_amb"]] <- list()
full_models[["EE_amb"]] <- list()

formulas <- list()

```


# Data

## Data view

Let's have a look at the original data. We remove several variables that, from a theortical consideration, are not relevant (and would sometimes cause collinearity) We also transform some values to a standard % format. 

```{r dataview}
#View(combined_analysis)
summary(combined_analysis)

combined_analysis_transformed <- combined_analysis %>% 
  mutate(
    pre_potential_EI = pre_potential_EI / 100)
  # ) %>% 
  # select(-subsidies_savings, - subsidies_REN, -percent_REN_economic, -percent_EE_economic, -percent_REN_planning, -percent_EE_planning)

summary(combined_analysis_transformed)

```


```{r translation tables, include=FALSE}
# Here we change the column names and prepare the dataframes. Assignements can now be adjusted in the Excel file.

all_variables_translation_table <- read_excel(path = "MLR_variables.xlsx")

# Select all variables in col_names_for_RE	col_names_for_EE, and prepare for renaming

all_variables <- list()
traslation_tables <- list()

# Select RE
all_variables[["RE"]] <- all_variables_translation_table %>% 
  filter(!is.na(col_names_for_RE)) %>% 
  select(col_name = col_names_for_RE, symbol)

traslation_tables[["RE"]] <- all_variables[["RE"]]$col_name
names(traslation_tables[["RE"]]) <- all_variables[["RE"]]$symbol

# Select EE
all_variables[["EE"]] <- all_variables_translation_table %>% 
  filter(!is.na(col_names_for_EE)) %>% 
  select(col_name = col_names_for_EE, symbol)

traslation_tables[["EE"]] <- all_variables[["EE"]]$col_name
names(traslation_tables[["EE"]]) <- all_variables[["EE"]]$symbol


## Gather and rename
                                           
combined_analysis_gathered <- combined_analysis_transformed %>% 
  select(-prioritization_scale, -Prio_EE_REN) %>% 
  gather(key = "variable", value = "value", one_of(all_variables$RE$col_name, all_variables$EE$col_name)) 

Ren_both_effects <- list()
Ren_both_effects[["long"]] <- combined_analysis_gathered %>% 
  filter(variable %in% c(all_variables$RE$col_name)) %>% 
  mutate(variable = fct_recode(variable, !!!traslation_tables$RE))

Ren_both_effects[["wide"]] <- Ren_both_effects[["long"]]  %>% 
  select(country_code, variable, value) %>% 
  spread(key = "variable", value = "value")

EE_both_effects <- list()
EE_both_effects[["long"]]  <- combined_analysis_gathered %>% 
  filter(variable %in% c(all_variables$EE$col_name)) %>% 
  mutate(variable = fct_recode(variable, !!!traslation_tables$EE))

EE_both_effects[["wide"]] <- EE_both_effects[["long"]]  %>% 
  select(country_code, variable, value) %>% 
  spread(key = "variable", value = "value")


all_variables_translation_table %>% 
  select(symbol, variable)

variable_names <- all_variables_translation_table$variable
names(variable_names) <- all_variables_translation_table$symbol


```

## Remove missing cases

First we need to remove missing cases.

```{r echo=FALSE}

#View(combined_analysis)

# Netherlands is missing

# From here on use wide format due to R methods.
 
EE_both_effects[["wide_no_na"]] <- na.omit(EE_both_effects[["wide"]])
paste("For EE data: Of", nrow(EE_both_effects$wide),"cases", nrow(EE_both_effects$wide_no_na),"are left due to missing data,")

# let's take this sample


Ren_both_effects[["wide_no_na"]] <- na.omit(Ren_both_effects[["wide"]]) %>% 
  filter(country_code %in% EE_both_effects[["wide_no_na"]]$country_code)

paste("For REN data: Of", nrow(Ren_both_effects$wide),"cases", nrow(Ren_both_effects$wide_no_na),"are left due to missing data,")


tibble(
  countries = EE_both_effects[["wide"]]$country_code,
  in_sample = EE_both_effects[["wide"]]$country_code %in% EE_both_effects[["wide_no_na"]]$country_code
) %>% 
  mutate(
    in_sample = if_else(in_sample, "In the finale sample", "Missing")
  ) %>% 
  tab_df()



# Subsample for checking the robustness of the model

ggplot(data = EE_both_effects[["wide_no_na"]], mapping = aes(x = S_size_services, y = EE_perf)) +
  geom_point() + 
  geom_smooth(method = "lm")
```


Some data is missing that we can't use in MLR, but most is there. Missing is percent_largest_generator for netherlands e.g..
TODO

## Marginal plots

Let's have a look at the data using marginal plots. This helps to see if any (potential) predictores are skewed. If this is the case, log() transformation will be applied in the regression model, which is a standard procedure to improve the model.

```{r marginal plot, echo=FALSE, message=FALSE, warning=FALSE}

Ren_both_effects %>%
  pluck("long") %>% 
  ggplot(mapping = aes(x = value)) +
    geom_histogram()+
    facet_wrap("variable", scales = "free") +
    labs(title = "Marginal plots for REN data")


EE_both_effects %>%
  pluck("long") %>% 
  ggplot(mapping = aes(x = value)) +
    geom_histogram()+
    facet_wrap("variable", scales = "free") +
    labs(title = "Marginal plots for EE data")

```

For the analysis of the RE data: P_ambition and S_coal are skewed to the right.

```{r variable transformation}



Ren_both_effects[["wide_no_na"]] <- Ren_both_effects[["wide_no_na"]]


EE_both_effects[["wide_no_na"]] <- EE_both_effects[["wide_no_na"]]

```

After transformation, it is still very skewed. Probably not good results if used. 
 
# 1. REN Performance

## Full model construction 

This model contains all possible IVs.

```{r}

formulas[["REN_perf_full"]] <- formula(RE_perf ~  
    P_density + P_ambition + P_REN +
    S_GDP + S_GAV + S_potential + S_effort + S_PPS + S_elec  + S_centr + S_size_industry + S_size_services)

full_models[["RE"]][["01_untransformed"]] <- lm(
  formula = formulas[["REN_perf_full"]], data = Ren_both_effects[["wide_no_na"]])


summary(full_models$RE$`01_untransformed`)

```


## Collinearity check

The selection of variables is essential. The preferred way is to have some theorethical knowledge from other studies. For a good model, there must no be any large collinearitie between variables.

```{r, warning=FALSE, echo=FALSE}
GGally::ggcorr(Ren_both_effects$wide_no_na, nbreaks = 5, label = TRUE, size = 2, label_size = 3, legend.position = "bottom", legend.size = 10)
```

All correlations above 0.8 are problematic. We don't find any. Let's also check this numerically based on the full (transfomred) model.

```{r}
vif(full_models$RE$`01_untransformed`)
```
VIF above 5 is critical and above 10 dangerous. S_GAV and S_effort are slightly collinear, but no critical VIFs. We do not need to exclude any of our IVs.

## Math variable selection

Let's apply a stepwise backward selectin of variables to the full model to improve things.

WARGNING: According to my stats script, "There is a general rule of the thumb, saying that there should be at least 5 times as many observations as the number of coefficients that are estimated.". I.e. based on the 8 IV, there should be 40 observations, however, we only have 25 observations, so effectively this type of selection may not yield a good result, and hence we will also perfom stepwise selections with other subsamples (see below).


```{r}
full_models[["RE"]][["03_selected"]] <- step(
  full_models[["RE"]][["01_untransformed"]], direction = "backward", trace = 0)
summary(full_models[["RE"]][["03_selected"]])

```

## Checking Model assumptions

It is very important that some assumptions are fulfilled. Otherwise, the findings may be wrong and can not be generalised. 


```{r}
resplot(full_models[["RE"]][["03_selected"]])
```

- Plot 1: The average error is probably 0 (within random resampled lines), which is good. There is no systematic error.
- Plot 2: Redsiduals are normall distributed, which is good 
- Plot 3: There seems to be heteroskedasity, but random resampling suggests that this line could have be gotten by chance.
- Plot 4: There aren't many influenctial cases. Good

All in all, statistical assumptions are fulfilled an the conclusions drawn from it are robust.

# 2. EE Performance 

## Full model construction 

This model contains all possible IVs.

```{r}

formulas[["EE_perf_full"]] <- formula(EE_perf ~  
    P_density + P_ambition + P_REN +
    S_GDP + S_GAV + S_potential + S_effort + S_PPS + S_elec + S_centr  + S_size_services + S_size_industry)

full_models[["EE"]][["01_untransformed"]] <- lm(
  formula = formulas[["EE_perf_full"]] , data = EE_both_effects[["wide_no_na"]])

summary(full_models$EE$`01_untransformed`)
```


## Collinearity check

```{r, warning=FALSE}
GGally::ggcorr(EE_both_effects$wide_no_na %>% select(-EE_perf), nbreaks = 5, label = TRUE, size = 2,label_size = 3, legend.position = "bottom", legend.size = 15)  +
  ggtitle ("Policy & socio-economic effects on EE progress")

```

```{r}
vif(full_models$EE$`01_untransformed`)
```

S_potential may be slightly collinear with S_GDP, however it's still below 10

## Math variable selection

Let's apply a stepwise backward selection of variables to the full model.


```{r}
full_models[["EE"]][["03_selected"]] <- step(full_models[["EE"]][["01_untransformed"]], direction = "backward", trace = 0)
summary(full_models[["EE"]][["03_selected"]])

```


## Checking Model assumptions

```{r}
resplot(full_models[["EE"]][["03_selected"]])
```

```{r}
# Check outlier
EE_both_effects[["wide_no_na"]] %>% 
  slice(c(14, 6))
```



# 3. REN Ambition

## Full model construction 

This model contains all possible IVs.

```{r}

full_models[["RE_amb"]][["01_untransformed"]] <- lm(
  P_ambition ~ 
    P_density + P_REN + 
    S_GDP + S_GAV + S_potential + S_effort + S_PPS + S_elec + S_centr  + S_size_services + S_size_industry, 
  data = Ren_both_effects[["wide_no_na"]])

summary(full_models$RE_amb$`01_untransformed`)
```


## Collinearity check

```{r, warning=FALSE}
GGally::ggcorr(Ren_both_effects$wide_no_na, nbreaks = 5, label = TRUE, size = 2,label_size = 3, legend.position = "bottom", legend.size = 15)

```

```{r}
vif(full_models$RE_amb$`01_untransformed`)
```

S_potential may be slightly collinear with S_effort, however it's still below 10. So not extremely problematic, but a bit if both variables will be in the final model.

## Math variable selection

Let's apply a stepwise backward selection of variables to the full model.


```{r}
full_models[["RE_amb"]][["03_selected"]] <- step(full_models[["RE_amb"]][["01_untransformed"]], direction = "backward", trace = 0)
summary(full_models[["RE_amb"]][["03_selected"]])

```


## Checking Model assumptions

```{r}
resplot(full_models[["RE_amb"]][["03_selected"]])
```

I'm not sure that we can trust this model. The Tukey-Anscombe plot is definitely skewed, despite being in the resampled area. Also there is a highly influential outlier (HR).

```{r}
# Check outlier
Ren_both_effects[["wide_no_na"]] %>% 
  slice(c(11))
```





# 4. REN Ambition

## Full model construction 

This model contains all possible IVs.

```{r}

full_models[["EE_amb"]][["01_untransformed"]] <- lm(
  P_ambition ~ 
    P_density + P_REN + 
    S_GDP + S_GAV + S_potential + S_effort + S_PPS + S_elec + S_centr  + S_size_services + S_size_industry, 
  data = EE_both_effects[["wide_no_na"]])

summary(full_models$EE_amb$`01_untransformed`)
```


## Collinearity check

```{r, warning=FALSE}
GGally::ggcorr(EE_both_effects$wide_no_na, nbreaks = 5, label = TRUE, size = 2,label_size = 3, legend.position = "bottom", legend.size = 15)

```

```{r}
vif(full_models$EE_amb$`01_untransformed`)
```

S_potential is higher than 10, this is problematic if it is in the final model


## Math variable selection
Let's apply a stepwise backward selection of variables to the full model.


```{r}
full_models[["EE_amb"]][["03_selected"]] <- step(full_models[["EE_amb"]][["01_untransformed"]], direction = "backward", trace = 0)
summary(full_models[["EE_amb"]][["03_selected"]])

```

S_potential not in the final model, so collinearity issue not a problem here.

## Checking Model assumptions

```{r}
resplot(full_models[["EE_amb"]][["03_selected"]])
```

These assumption plots look better. No highly influential outliers

# Overview tables

```{r}
tab_model(
  full_models$RE$`03_selected`, full_models$EE$`03_selected`,
  dv.labels = c("RE performance", "EE performance"),
  pred.labels = variable_names,
  title = "Explaining REN Performance",
  p.style = "stars",
  show.std = TRUE,
  p.threshold = c(0.1, 0.05, 0.01),
  digits = 3)
```


```{r}
tab_model(
  full_models$RE_amb$`03_selected`, full_models$EE_amb$`03_selected`,
  dv.labels = c("RE ambition", "EE ambition"),
  pred.labels = variable_names,
  title = "Explaining REN Ambition",
  p.style = "stars",
  show.std = TRUE,
  p.threshold = c(0.1, 0.05, 0.01),
  digits = 3)
```